# Report of Comorbidity Project Pairs analysis

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  metric_table <- data[['table_metrics.txt']]

  HyI_distribution_plot <- function(df, column, log, xlab, title){
    plt <- ggplot(df, aes(x=get(column))) + geom_density(alpha=0.4) 
    plt <- plt + geom_vline(aes(xintercept = 2, colour = "red"))
    plt <- plt + xlab(xlab)
    plt <- plt + ylab("Density")
    plt <- plt + theme(legend.position="none")
    
    if(log == 10){
        plt <- plt + scale_x_log10()
    }else if (log == 2){
        plt <- plt + scale_x_continuous(trans='log2')
    }
    plt <- plt + ggtitle (title)
    return(plt)
}

pairs_prevalence_graph <- function (df, title, log, xlab){
  colnames(df) <- c("HPO-A", "HPO-B", "Prevalence-A", "Prevalence-B", "Type")
  df <- mutate(df, prevalence_mean=(rowMeans(df[c("Prevalence-A", "Prevalence-B")], na.rm=TRUE))) 

  df_1 <- rbind(filter(df, Type == "more_spec"), filter(df, Type == "less_spec"), filter(df, Type == "unconnected"))

  plt <- ggplot(df_1, aes(x=prevalence_mean, colour=Type)) + geom_line(aes(linetype=Type), stat="density")
  plt <- plt + ggtitle(title)
  plt <- plt + theme(legend.title=element_blank())
  plt <- plt + xlab(xlab)
  plt <- plt + ylab("Density")

  if(log == TRUE){
    plt <- plt + scale_x_log10()
    }

  return(plt)
}


hpo_prevalence_graph <- function(df, title, xlab){
  
  colnames(df) <- c("HPO", "Prevalence")

  plt <- ggplot(df, aes(x=Prevalence)) + geom_line(stat="density") 
  plt <- plt + ggtitle(title)
  plt <- plt + ylab("Density")
  plt <- plt + xlab(xlab)
  plt <- plt + scale_x_log10() + scale_y_continuous()
  

  

  return(plt)
}

  produce_barplot <- function( column, title, ylab, second_column=NULL, add=FALSE){
    df <- summarySE(metric_table, measurevar=column, groupvars=c("Type"))
    position <- c("more_spec", "more_spec_l_rdm", "more_spec_n_rdm", "less_spec", "less_spec_l_rdm", "less_spec_n_rdm", "unconnected")

    plt <- ggplot(df, aes(x=Type, y=get(column), fill=Type)) + scale_x_discrete(limits = position)
    plt <- plt + geom_bar(position=position_dodge(), stat="identity")
    if(!is.null(second_column)){
      second_data_frame <- summarySE(metric_table, measurevar=second_column, groupvars=c("Type"))
      plt <- plt + geom_point(mapping = aes(x = Type, y = second_data_frame[[second_column]])) 
    }
    plt <- plt + geom_errorbar(aes(ymin=get(column)-sd, ymax=get(column)+sd),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) #+
    if(add){
      #plt <- plt + theme(plot.margin = unit(c(1,5,-30,6),units="points")
      plt <- plt + theme(axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          axis.ticks.x=element_blank()  
        )
    }else{
      plt <- plt + theme(axis.text.x = element_text(angle = 45, hjust = 1), 
                      axis.title.x=element_blank()
                      )
    }
    plt <- plt + ylab(ylab) #+
    if(title != ''){
      plt <- plt + ggtitle(title) #+
    }
    plt <- plt + theme(legend.position="none") #+
    
    return(plt)
  }

  produce_stacked_barplot <- function( column1, column2, column3, title, ylab, add=FALSE, legend=FALSE){
    library(dplyr)
    library(reshape)
    
    df <- select(metric_table, column1, column2, column3)
    new_df <- melt(df)
    position <- c("more_spec", "more_spec_l_rdm", "more_spec_n_rdm", "less_spec", "less_spec_l_rdm", "less_spec_n_rdm", "unconnected")
      
    plt <- ggplot(new_df, aes(x=Type, y=value, fill=variable)) + scale_x_discrete(limits = position) 
    plt <- plt + geom_bar(stat='identity', position="fill")

    plt <- plt + theme(axis.text.x = element_text(angle = 45, hjust = 1), 
                      axis.title.x=element_blank()
                      )
    if(add){
      #plt <- plt + theme(plot.margin = unit(c(1,5,-30,6),units="points"))
      plt <- plt + theme(legend.position="none")
      
    }else{
      plt <- plt + scale_fill_discrete(labels = c("Consistent GO terms", "Emergent GO terms"))
      plt <- plt + theme(legend.position="bottom", legend.direction="vertical",
                 legend.title = element_blank())
    }
    plt <- plt + ylab(ylab) #+
    if(title != ''){
      plt <- plt + ggtitle(title) #+
    }

    return(plt)
  }

    load_files <- function(data, column_names, path_column, header){
        factor_columns <- match(column_names, names(data))
        factor_combinations <- unique(data[column_names])
        all_data <- data.frame()
        for(row in 1:nrow(factor_combinations)){
                combination <- as.vector(t((factor_combinations[row,]))) #extract row AND convert to vector
                check_combination <- data[factor_columns] == combination[col(data[factor_columns])]
                name_list <- paste(combination, collapse='_')
                paths <- data[[path_column]][which(apply(check_combination, 1, sum) == length(combination))]
                count = 1

                for(file_path in paths){
                        files <- read.table(file_path, sep="\t", header=header)
                        all_data <- rbind(all_data, files)   
                }             
        }
       
        return(all_data)

      }


  load_comention_files <- function(data, column_names, path_column, header, col_name_vector, group1, group2, group3){
        factor_columns <- match(column_names, names(data))
        factor_combinations <- unique(data[column_names])
        all_data <- data.frame()
        for(row in 1:nrow(factor_combinations)){
                combination <- as.vector(t((factor_combinations[row,]))) #extract row AND convert to vector
                check_combination <- data[factor_columns] == combination[col(data[factor_columns])]
                name_list <- paste(combination, collapse='_')
                paths <- data[[path_column]][which(apply(check_combination, 1, sum) == length(combination))]
                count = 1

                for(file_path in paths){
                        files <- read.table(file_path, sep="\t", header=header)
                        files$V4 <- NULL
                        files <- cbind(files ,rep(c(paste(name_list)), nrow(files)))
                        files <- cbind(files ,rep(c(paste(name_list,count, sep="_")), nrow(files)))
                        if(startsWith(name_list, group1)){
                          files <- cbind(files ,rep(c(paste(group1)), nrow(files)))
                        }else if(startsWith(name_list, group2)){
                          files <- cbind(files ,rep(c(paste(group2)), nrow(files)))
                        }else{
                          files <- cbind(files ,rep(c(paste(group3)), nrow(files)))
                        }
                        colnames(files) <- col_name_vector
                        count <- count + 1

                        all_data <- rbind(all_data, files)   
                }             
        }
       
        return(all_data)

      }


      produce_density_plot <- function(data_table, values, net_names, net_types, x_scale = FALSE, x_lab, title, facet = FALSE){
        plt <- ggplot(data_table, aes(x=get(values))) + geom_density(aes(group=get(net_names), colour=get(net_types)))
        col <- c("firebrick1", "orange4", "deepskyblue2", "lightseagreen")

        if (facet == TRUE){
          plt <- plt + facet_wrap( ~ groups, scales = "free")
        }
        if(x_scale == TRUE){
          plt <- plt + scale_x_continuous(limits = c(0, 1))
        }
        plt <- plt + ylim(0, 4.5)
        plt <- plt + xlab(x_lab)
        plt <- plt + ylab("Density")
        plt <- plt + ggtitle(title)
        plt <- plt + theme(legend.title=element_blank())
        plt <- plt + scale_fill_manual(values = col)
        plt <- plt + scale_color_manual(values = col)

        return(plt)
      }


```


We want to study comorbidity relationships between phenotypes pairs from DECIPHER database.

We have built different networks:  
  
**1. more_spec (Comorbidity Network)**: HPO pairs with a HyI >= 2.  
**2. less_spec**: HPO pairs with a HyI < 2.  
**4. random_links**: randmly shufflerd the associations between HPO pairs, while keeping the number of edges and nodes unchanged.  
**5. random_nodes**: Have de same properties and topological structure than the Hypergeometic_phen_reliable_terms. But the number of edges per node change.  
**6. unconnected**: Random selection of HPO pairs that are not connected by any patient in DECIPHER database (negative control). 


### Properties of the Phenotype-Phenotype pairs lists and random models


```{r  echo=FALSE, warning=FALSE, message=FALSE}
  build_metric_table <- read.table("results/build_metrics", sep="\t", header = FALSE)
  colnames(build_metric_table) <- c("Metric", "Value")
  kable(build_metric_table)
```
  
**Table 1**: Properties of the phenotype-phenotype pairs lists and random models.

  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)

hyi_data <- read.table("results/phen2phen_net", header=FALSE)
colnames(hyi_data) <- c("HPO-A", "HPO-B", "HyI")
HyI_distribution_plot(hyi_data, "HyI",log = FALSE, "Hypergeometric Index", "Hypergeometric Index Distribution")

single_prevalence_data <- read.table("results/all_hpo_prevalence", header=FALSE)
hpo_prevalence_graph(single_prevalence_data, "Phenotype Prevalence in DECIPHER", "Prevalence (%)")


prevalence_pairs_data <- load_files(metric_table, "Type", "decipher_pairs_prevalence", header=FALSE)
pairs_prevalence_graph(prevalence_pairs_data, "Phenotype Pairs Prevalence in DECIPHER", log = TRUE, "Mean Prevalence (%)")
  


```
  
**Figure 1**: Properties of the Phenotype-Phenotype pairs  


### Number of genes mapping to each phenotype pair

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  gene_table <- load_files(metric_table, "Type", "pairs_genes_table", header=FALSE)
  colnames(gene_table) <- c("HPO-A", "HPO-B", "GENES-A", "GENES-B", "Union_genes", "Intersection_genes", "Net_type")

  kable(ddply(gene_table, .(Net_type), summarize, "Genes mean per pair" = mean(Union_genes)))


```

### Topological Analysis

#### MEANS
```{r  echo=FALSE, warning=FALSE, message=FALSE}
	metric_table <- data[['table_metrics.txt']]
	kable(ddply(metric_table, .(Type), summarize, Clustering.coef= mean(Clustering.coef), Diameter = mean(Diameter), Average.minimum.path= mean(Average.minimum.path), Number.of.nodes = mean(Number.of.nodes)))  

```

#### STANDARD DEVIATION

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  metric_table <- data[['table_metrics.txt']]
  kable(ddply(metric_table, .(Type), summarize, Clustering.coef= sd(Clustering.coef), Diameter = sd(Diameter), Average.minimum.path= sd(Average.minimum.path), Number.of.nodes = sd(Number.of.nodes)))  

```  
  
**Table 2**: Topological Analysis results from the different Pairs Lists. 

### Overlap of the pairs lists with known diseases

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  grid.arrange(  
    produce_barplot("confirmed_pairs_OMIM_cut_1", "At least one Disease", "Pairs", add=TRUE),
    produce_barplot("confirmed_pairs_ORPHA_cut_1", "At least one Disease", "Pairs", add=TRUE),
    produce_barplot("confirmed_pairs_OMIM_cut_2", "At least two Diseases", "Pairs", add=TRUE),
    produce_barplot("confirmed_pairs_ORPHA_cut_2", "At least two Diseases", "Pairs", add=TRUE),
    ncol= 2, nrow = 2,
    top = "OMIM / ORPHANET"
  )
  grid.arrange(  
    produce_barplot("confirmed_pairs_OMIM_cut_3", "At least three Diseases", "Pairs"),
    produce_barplot("confirmed_pairs_ORPHA_cut_3", "At least three Diseases", "Pairs"),
    ncol= 2
  )
```
  
**Figure 2**: Number of phenotype pairs that appear together in OMIM (left) and Orphanet (right) diseases. 

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  metric_table <- data[['table_metrics.txt']]
  kable(ddply(metric_table, .(Type), summarize, confirmed_pairs_OMIM= mean(confirmed_pairs_OMIM_cut_3), confirmed_pairs_ORPHA= mean(confirmed_pairs_ORPHA_cut_3), Union_Diseases = mean(Union_Diseases_cut_3), Intersection_Diseases = mean(Intersection_Diseases_cut_3)))  

```
  
**Table 3**: Overlap with diseases summary. 
  

## Comention Analysis.
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}

comention_data <- load_comention_files(metric_table, 'Type', 'comention_analysis', FALSE, c("HPOA", "HPOB", "Pvalue", "Net_type", "Net_name", "groups"), "more_spec", "less_spec", "unconnected")
comention_suitable <- subset(comention_data, Net_type %in% c("more_spec", "more_spec_l_rdm", "less_spec", "less_spec_l_rdm"))

produce_density_plot(comention_suitable, "Pvalue", "Net_name", "Net_type", x_scale = TRUE, "P-value", "P-value distribution", facet = TRUE)


```
  
**Figure 3**: Comention pvalue density plot.
  

## PMID per Pairs list.
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}

pmid_data <- load_comention_files(metric_table, 'Type', 'pmid_analysis', FALSE, c("HPOA", "HPOB", "Union_PMID", "Net_type", "Net_name", "groups"), "more_spec", "less_spec", "unconnected")   
pmid_suitable <- subset(pmid_data, Net_type %in% c("more_spec", "more_spec_l_rdm", "less_spec", "less_spec_l_rdm"))

kable(ddply(pmid_suitable, .(Net_type), summarize, Union_PMID_mean= mean(Union_PMID), Union_PMID_sd= sd(Union_PMID), Union_PMID_median = median(Union_PMID), Union_PMID_IQR = IQR(Union_PMID))) 
```
  
**Table 4**: Union PMID metrics per network type.
  


### Emergent and Consistent functional systems
  
#### KEGG   
```{r  echo=FALSE, warning=FALSE, message=FALSE}
  produce_barplot("hpo_pairs_with_at_least_1term_kegg_0.05", "Pairs number with KEGG enrichment", "Pairs", add=FALSE)

  grid.arrange(
    produce_stacked_barplot("Type", "pairs_with_consistent_terms_kegg_0.05", "pairs_with_emergent_terms_kegg_0.05", "Pairs Consistent/emergent pathways", "Pairs", add = TRUE),
    produce_stacked_barplot("Type", "consistent_enriched_terms_kegg_0.05", "emergent_enriched_terms_kegg_0.05", "Consistent/Emergent pathways", "Pathways", add=FALSE),
    ncol=3
  )
  
```
  
**Figure 4**: KEGG Enrichment Analysis. **Top**: Number of pairs with at least one consistent or emergent pathways. **Bottom left**: Number of Consistent/Emergent Pathways. **Bottom right**: Number of Pairs with Consistent/Emergent Pathways.    
  

#### Mean
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
kable(ddply(metric_table, .(Type), summarize, "HPO pairs with at least 1 term KEGG" = mean(hpo_pairs_with_at_least_1term_kegg_0.05), "Consistent enriched terms KEGG"= mean(consistent_enriched_terms_kegg_0.05), "Emergent enriched terms KEGG" = mean(emergent_enriched_terms_kegg_0.05), "Pairs with consistent terms KEGG"=mean(pairs_with_consistent_terms_kegg_0.05), "Pairs with emergent terms KEGG"=mean(pairs_with_emergent_terms_kegg_0.05)))
```
  
#### SD
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
kable(ddply(metric_table, .(Type), summarize, "HPO pairs with at least 1 term KEGG" = sd(hpo_pairs_with_at_least_1term_kegg_0.05), "Consistent enriched terms KEGG"= sd(consistent_enriched_terms_kegg_0.05), "Emergent enriched terms KEGG" = sd(emergent_enriched_terms_kegg_0.05), "Pairs with consistent terms KEGG"=sd(pairs_with_consistent_terms_kegg_0.05), "Pairs with emergent terms KEGG"=sd(pairs_with_emergent_terms_kegg_0.05)))
```

**Table 5**: Summary of KEGG functional analysis
  
#### REACTOME

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  produce_barplot("hpo_pairs_with_at_least_1term_reactome_0.05", "Pairs number with REACTOME enrichment", "Pairs", add=FALSE)
  
  grid.arrange(
    produce_stacked_barplot("Type", "pairs_with_consistent_terms_reactome_0.05", "pairs_with_emergent_terms_reactome_0.05", "Pairs Consistent/emergent pathways", "Pairs", add = TRUE),
    produce_stacked_barplot("Type", "consistent_enriched_terms_reactome_0.05", "emergent_enriched_terms_reactome_0.05", "Consistent/Emergent pathways", "Pathways", add=FALSE),
    ncol=2
  )
  
```  
  
**Figure 5**: Reactome Enrichment Analysis. **Top**: Number of pairs with at least one consistent or emergent pathways. **Bottom left**: Number of Consistent/Emergent Pathways. **Bottom right**: Number of Pairs with Consistent/Emergent Pathways.
  
#### Mean
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
kable(ddply(metric_table, .(Type), summarize, "HPO pairs with at least 1 term Reactome" = mean(hpo_pairs_with_at_least_1term_reactome_0.05), "Consistent enriched terms Reactome" = mean(consistent_enriched_terms_reactome_0.05), "Emergent enriched terms Reactome" = mean(emergent_enriched_terms_reactome_0.05), "Pairs with consistent terms Reactome"=mean(pairs_with_consistent_terms_reactome_0.05), "Pairs with emergent terms Reactome"=mean(pairs_with_emergent_terms_reactome_0.05)))
```  
  
#### SD
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
kable(ddply(metric_table, .(Type), summarize, "HPO pairs with at least 1 term Reactome" = sd(hpo_pairs_with_at_least_1term_reactome_0.05), "Consistent enriched terms Reactome" = sd(consistent_enriched_terms_reactome_0.05), "Emergent enriched terms Reactome" = sd(emergent_enriched_terms_reactome_0.05), "Pairs with consistent terms Reactome"=sd(pairs_with_consistent_terms_reactome_0.05), "Pairs with emergent terms Reactome"=sd(pairs_with_emergent_terms_reactome_0.05)))
```
  
**Table 6**: Summary of Reactome functional analysis

#### GO  

```{r  echo=FALSE, warning=FALSE, message=FALSE}
  produce_barplot("hpo_pairs_with_at_least_1term_go_0.05", "Pairs number with GO enrichment", "Pairs", add=FALSE)
  
  grid.arrange(
    produce_stacked_barplot("Type", "pairs_with_consistent_terms_go_0.05", "pairs_with_emergent_terms_go_0.05", "Pairs Consistent/emergent pathways", "Pairs", add = TRUE),
    produce_stacked_barplot("Type", "consistent_enriched_terms_go_0.05", "emergent_enriched_terms_go_0.05", "Consistent/Emergent Pathways", "Pathways", add=FALSE),
    ncol=2  
  )
  
```
  
**Figure 6**: GO Enrichment Analysis. **Top**: Number of pairs with at least one consistent or emergent pathways. **Bottom left**: Number of Consistent/Emergent Pathways. **Bottom right**: Number of Pairs with Consistent/Emergent Pathways.
  
#### Mean

```{r  echo=FALSE, warning=FALSE, message=FALSE}
kable(ddply(metric_table, .(Type), summarize, "HPO pairs with at least 1 term GO" = mean(hpo_pairs_with_at_least_1term_go_0.05), "Consistent enriched terms GO"= mean(consistent_enriched_terms_go_0.05), "Emergent enriched terms GO" = mean(emergent_enriched_terms_go_0.05), "Pairs with consistent terms GO"=mean(pairs_with_consistent_terms_go_0.05), "Pairs with emergent terms GO"=mean(pairs_with_emergent_terms_go_0.05)))
```
  
#### SD
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
kable(ddply(metric_table, .(Type), summarize, "HPO pairs with at least 1 term GO" = sd(hpo_pairs_with_at_least_1term_go_0.05), "Consistent enriched terms GO"= sd(consistent_enriched_terms_go_0.05), "Emergent enriched terms GO" = sd(emergent_enriched_terms_go_0.05), "Pairs with consistent terms GO"=sd(pairs_with_consistent_terms_go_0.05), "Pairs with emergent terms GO"=sd(pairs_with_emergent_terms_go_0.05)))
```
  
**Table 7**: Summary of GO functional analysis


