# Report 6: Patient Details
  
```{r  echo=FALSE, warning=FALSE, message=FALSE}
  metric_table <- data[['more-spec_table_metrics.txt']]
  library(dplyr)
  

#Load file from a metric file
  load_files <- function(data, column_names, path_column, header){
        factor_columns <- match(column_names, names(data))
        factor_combinations <- unique(data[column_names])
        
        for(row in 1:nrow(factor_combinations)){
                combination <- as.vector(t((factor_combinations[row,]))) #extract row AND convert to vector
                check_combination <- data[factor_columns] == combination[col(data[factor_columns])]
                paths <- data[[path_column]][which(apply(check_combination, 1, sum) == length(combination))]
                
                for (file_path in paths){
                  files <- read.table(file_path, sep="\t", header=header, quote="", stringsAsFactors = FALSE)
                    
                  }
                
                }
           return(files)     
        }

#Function which group different dataframes by patients   
    group_patient_information <- function(df1, df2, df3, df4){
      patients <- unique(df1$Patient)
      all_patients <- list()

      for (i in patients){
        all_groups <- list()
        summary_groups <- filter(df1, Patient == i)
        if (nrow(summary_groups) != 0){
          all_groups[["sumary"]] <- summary_groups
        }
        gene_groups <- filter(df4, Patient == i)
        if (nrow(gene_groups) != 0){
          all_groups[["gene"]] <- gene_groups
        }

        hp_groups <- filter(df2, Patient == i)
        if (nrow(hp_groups) != 0){
          all_groups[["hpo"]] <- hp_groups
        }

        system_groups <- filter(df3, Patient == i)
        if (nrow(system_groups) != 0){
          all_groups[["system"]] <- system_groups
          phenotypes_vector_length <- c()
          phenotypes <- list()
          for(patient in 1:length(system_groups$Shared_hpos)){
            phenotypes[patient] <- strsplit(system_groups$Shared_hpos[patient], ", ")
            number_hpo <- strsplit(system_groups$Shared_hpos[patient], ", ")[[1]]
            phenotypes_vector_length[patient] <- length(number_hpo)
          }
          unique_hpo_list <- unique(unlist(phenotypes))
          all_groups[["phenotype_length"]] <- length(unique_hpo_list)
          
          systems_index <- with(system_groups, order(phenotypes_vector_length, decreasing = TRUE))
          systems_groups_ordered <- system_groups[systems_index,]
          rownames(systems_groups_ordered) <- NULL
          all_groups[["system"]] <- systems_groups_ordered

        }else{
          all_groups[["phenotype_length"]] <- 0
        }

        name <- paste0("Patient_", i)
        all_patients[[name]] <- all_groups 
        }
        return(all_patients)  
    }

  produce_scatterplot <- function(data_set, x_axis, y_axis, x_lab, y_lab, add=FALSE){

      plt <- ggplot(data_set, aes(x=get(x_axis), y=get(y_axis)))

      if (add){
        plt <- plt + geom_count(stat = "sum", show.legend = TRUE)
        plt <- plt + theme(legend.title=element_blank())
        plt <- plt + scale_size(name   = range(1:150),
               breaks = c(1, 50, 100, 150),
               labels = c(1, 50, 100, 150))
      } else {
        plt <- plt + geom_point()
      }

      plt <- plt + xlab(x_lab)
      plt <- plt + ylab(y_lab)
      plt <- plt + theme(legend.title=element_blank())

      return(plt)
    }

  get_specific_go_terms <- function(df){
    library(GOSim)
    library(GO.db)
    
    #Get child terms
    setOntology(ont = "BP")
    go_l <- getChildren()

    #load final dataframe
    df1 <- data_frame()
    
    #Get patient list
    patient_l <- unique(df$Patient)
    
    #For each patient get the overlapping clusters
    for (patient in patient_l){
      patient_groups <- filter(df, Patient == patient)  
      clusters_l <- unique(patient_groups$Cluster)
      
      #For each cluster, get the overlapping systems
      for (cluster in clusters_l){
        clusters_groups <- filter(patient_groups, Cluster == cluster)
        system_l <- unique(clusters_groups$Systems)
        child_l <- c()
        
        #If there are more than one system get the child terms of each system and load them in a list
        if (length(system_l) > 1){
          for(system in system_l){
            child_l <- c(child_l, go_l[[system]])
          }
          #If a child term is the list of system in the cluster, eliminate the line in the df. Finally store the final df in the final df
          for (child in child_l){
            if(child %in% system_l){
              clusters_groups <- (clusters_groups[!clusters_groups$Systems == child, ])
            }
          }
          df1 <- rbind(df1, clusters_groups)
        #If the cluster only have one system load the df to the final df
        }else{
          df1 <- rbind(df1, clusters_groups)
        }
        
      }
    }
    return(df1)
  }

```

## Summary tables

```{r  echo=FALSE, warning=FALSE, message=FALSE}
df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= mean(patient_overlaping_with_clusters), "Number of patients overlap with functional coherent clusters"= mean(patient_overlaping_with_clusters_functional_coherent))

kable(t(df))

```

### Patients that overlap with clusters build with GO valid pairs 


```{r  echo=FALSE, warning=FALSE, message=FALSE}

go_df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= number_of_patients_corresponding_to_at_least_1_cluster_go_0.05, "Number of patients overlap with go functional coherent and OMIM clusters"= number_of_patients_corresponding_both_funsys_omim_go_0.05, "Number of patients overlap with go functional coherent clusters" = number_of_patients_corresponding_to_funsys_go_0.05, "Number of patients orvelap with OMIM clusters" = number_of_patients_corresponding_to_omim_go_0.05)

kable(t(go_df))

```

### Patients that overlap with clusters build with KEGG valid pairs 


```{r  echo=FALSE, warning=FALSE, message=FALSE}

kegg_df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= number_of_patients_corresponding_to_at_least_1_cluster_kegg_0.05, "Number of patients overlap with kegg functional coherent and OMIM clusters"= number_of_patients_corresponding_both_funsys_omim_kegg_0.05, "Number of patients overlap with kegg functional coherent clusters" = number_of_patients_corresponding_to_funsys_kegg_0.05, "Number of patients orvelap with OMIM clusters" = number_of_patients_corresponding_to_omim_kegg_0.05)

kable(t(kegg_df))

```

### Patients that overlap with clusters build with Reactome valid pairs 


```{r  echo=FALSE, warning=FALSE, message=FALSE}

reactome_df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= number_of_patients_corresponding_to_at_least_1_cluster_reactome_0.05, "Number of patients overlap with reactome functional coherent and OMIM clusters"= number_of_patients_corresponding_both_funsys_omim_reactome_0.05, "Number of patients overlap with reactome functional coherent clusters" = number_of_patients_corresponding_to_funsys_reactome_0.05, "Number of patients orvelap with OMIM clusters" = number_of_patients_corresponding_to_omim_reactome_0.05)

kable(t(reactome_df))

```

## Number of cluster vs number of HPO per patient  


```{r  echo=FALSE, warning=FALSE, message=FALSE, dpi=600, res=500}
  
  all_data_table <- load_files(metric_table, 'Type', 'patients_systems_cases', TRUE)
  produce_scatterplot(all_data_table, "Prof_length", "Nº_coincedence_clusters", "Number of HPO per patient", "Number of clusters", add=TRUE)

```

**Figure 1**: ***Number of cluster related to the number of HPO per patient*** (*x-axis Number of HPO per patient. y-axis  Number of cluster containing some HPO patient profile*). As we can see, the profile of the patients point to different clusters.



## Example of Patients Data

```{r  echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
  library(kableExtra)

#Load files and add hyperlinks
  
  hpo <- load_files(metric_table, 'Type', 'patient_hpo_coincident_with_clusters', FALSE)
  colnames(hpo) <- c("Patient", "HPO", "Name")

  summary <- load_files(metric_table, 'Type', 'patient_overlap_summary', TRUE)

  genes <- load_files(metric_table, 'Type', 'pat2gene', FALSE)
  colnames(genes) <- c("Patient", "Gene list from patient CNV")

  go_df <- load_files(metric_table, 'Type', 'patient_coincidence_with_clusters_and_gene_systems_diseases_go_0.05', TRUE)
  go <- get_specific_go_terms(go_df)
  if(nrow(go) > 0){  
    for (i in 1:length(go$Systems)){
      url <- paste0("http://amigo.geneontology.org/amigo/term/", go$Systems[i])
      go$Systems[i] <- text_spec(go$Systems[i], link = url)
    }
  }
  
  reactome <- load_files(metric_table, 'Type', 'patient_coincidence_with_clusters_and_gene_systems_diseases_reactome_0.05', TRUE)
  if(nrow(reactome) > 0){  
    for (i in 1:length(reactome$Systems)){
      url <- paste0("https://reactome.org/content/detail/", reactome$Systems[i])
      reactome$Systems[i] <- text_spec(reactome$Systems[i], link = url)
    }
  }

  kegg <- load_files(metric_table, 'Type', 'patient_coincidence_with_clusters_and_gene_systems_diseases_kegg_0.05', TRUE)
  if(nrow(kegg) > 0){
    for (i in 1:length(kegg$Systems)){
      url <- paste0("https://www.genome.jp/dbget-bin/www_bget?pathway+", kegg$Systems[i])
      kegg$Systems[i] <- text_spec(kegg$Systems[i], link = url)
    }
  }
  

  systems <- rbind(go, reactome, kegg)
  
  
#Group by patients

all_patients <- group_patient_information(summary, hpo, systems, genes)

#Order by phenotype overlap 
length_vector <- c()
  for(i in 1:length(all_patients)){
    list <- all_patients[[i]]
    phenotype_length <- list["phenotype_length"]
    length_vector[i] <- phenotype_length
    }

ordered_patients <- all_patients[order(unlist(length_vector), decreasing=TRUE)]  

#Print data
for(i in 1:length(ordered_patients)){
  list <- ordered_patients[[i]]
  if (length(list) == 5){
    cat("_ _ _", "\n", "_ _ _", "\n", "_ _ _", "\n")
    patient_name <- unlist(strsplit(names(ordered_patients[i]), "_"))
    cat("# ", patient_name, "\n")
  
    list$phenotype_length <- NULL
    for(i in list){
      print(knitr::kable(i, format="markdown"))
      cat("\n")
    }  
  }
  
}

    


  

```