```{r  echo=FALSE, warning=FALSE, message=FALSE}
  metric_table <- data[['more-spec_table_metrics.txt']]
  library(dplyr)
  


  load_files <- function(data, column_names, path_column, header){
        factor_columns <- match(column_names, names(data))
        factor_combinations <- unique(data[column_names])
        
        for(row in 1:nrow(factor_combinations)){
                combination <- as.vector(t((factor_combinations[row,]))) #extract row AND convert to vector
                check_combination <- data[factor_columns] == combination[col(data[factor_columns])]
                paths <- data[[path_column]][which(apply(check_combination, 1, sum) == length(combination))]
                
                for (file_path in paths){
                  files <- read.table(file_path, sep="\t", header=header, quote="", stringsAsFactors = FALSE)
                    
                  }
                
                }
           return(files)     
        }
   
    group_cluster_information <- function(df1, df2){
      clusters <- unique(df1$Patient)
      all_clusters <- list()

      for (i in clusters){
        all_groups <- list()
        hp_groups <- filter(df1, Patient == i)
        if (nrow(hp_groups) != 0){
          all_groups[["hpo"]] <- hp_groups
        }

        system_groups <- filter(df2, Patient == i)
        if (nrow(system_groups) != 0){
          all_groups[["system"]] <- system_groups
        }

        name <- paste0("Patient_", i)
        all_clusters[[name]] <- all_groups 
        }
        return(all_clusters)  
    }

  produce_scatterplot <- function(data_set, x_axis, y_axis, x_lab, y_lab, add=FALSE){

      plt <- ggplot(data_set, aes(x=get(x_axis), y=get(y_axis)))

      if (add){
        plt <- plt + geom_count(stat = "sum", show.legend = TRUE)
        plt <- plt + theme(legend.title=element_blank())
        plt <- plt + scale_size(name   = range(1:150),
               breaks = c(1, 50, 100, 150),
               labels = c(1, 50, 100, 150))
      } else {
        plt <- plt + geom_point()
      }

      plt <- plt + xlab(x_lab)
      plt <- plt + ylab(y_lab)

      return(plt)
    }

```

## Summary tables

```{r  echo=FALSE, warning=FALSE, message=FALSE}
df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= mean(patient_overlaping_with_clusters), "Number of patients overlap with functional coherent clusters"= mean(patient_overlaping_with_clusters_functional_coherent))

kable(t(df))

```

### Patients that overlap with clusters build with GO valid pairs 


```{r  echo=FALSE, warning=FALSE, message=FALSE}

go_df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= number_of_patients_corresponding_to_at_least_1_cluster_go_0.05, "Number of patients overlap with go functional coherent and OMIM clusters"= number_of_patients_corresponding_both_funsys_omim_go_0.05, "Number of patients overlap with go functional coherent clusters" = number_of_patients_corresponding_to_funsys_go_0.05, "Number of patients orvelap with OMIM clusters" = number_of_patients_corresponding_to_omim_go_0.05)

kable(t(go_df))

```

### Patients that overlap with clusters build with KEGG valid pairs 


```{r  echo=FALSE, warning=FALSE, message=FALSE}

kegg_df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= number_of_patients_corresponding_to_at_least_1_cluster_kegg_0.05, "Number of patients overlap with kegg functional coherent and OMIM clusters"= number_of_patients_corresponding_both_funsys_omim_kegg_0.05, "Number of patients overlap with kegg functional coherent clusters" = number_of_patients_corresponding_to_funsys_kegg_0.05, "Number of patients orvelap with OMIM clusters" = number_of_patients_corresponding_to_omim_kegg_0.05)

kable(t(kegg_df))

```

### Patients that overlap with clusters build with Reactome valid pairs 


```{r  echo=FALSE, warning=FALSE, message=FALSE}

reactome_df <- ddply(metric_table, .(Type), summarize,  "Number of patients to clusters"= number_of_patients_corresponding_to_at_least_1_cluster_reactome_0.05, "Number of patients overlap with reactome functional coherent and OMIM clusters"= number_of_patients_corresponding_both_funsys_omim_reactome_0.05, "Number of patients overlap with reactome functional coherent clusters" = number_of_patients_corresponding_to_funsys_reactome_0.05, "Number of patients orvelap with OMIM clusters" = number_of_patients_corresponding_to_omim_reactome_0.05)

kable(t(reactome_df))

```

## Number of cluster vs Number of HPO per patient  


```{r  echo=FALSE, warning=FALSE, message=FALSE}
  
  all_data_table <- load_files(metric_table, 'Type', 'patients_systems_cases', TRUE)
  produce_scatterplot(all_data_table, "Prof_length", "Nº_coincedence_clusters", "Number of HPO per patient", "Number of Clusters", add=TRUE)

```

**Figure 1**: ***Number of cluster related to the number of HPO per patient*** (*x-axis Number of HPO per patient. y-axis  Number of cluster containing some HPO patient profile*). As we can see, the profile of the patients point to different clusters.



## Patient Data

```{r  echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
  library(kableExtra)
  
  hpo <- load_files(metric_table, 'Type', 'patient_hpo_genes_go_0.05', TRUE)

  go <- load_files(metric_table, 'Type', 'patient_coincidence_with_clusters_and_gene_systems_go_0.05', TRUE)
  for (i in 1:length(go$Systems)){
    url <- paste0("http://amigo.geneontology.org/amigo/term/", go$Systems[i])
    go$Systems[i] <- text_spec(go$Systems[i], link = url)
  }

  kegg <- load_files(metric_table, 'Type', 'patient_coincidence_with_clusters_and_gene_systems_kegg_0.05', TRUE)
  for (i in 1:length(kegg$Systems)){
    url <- paste0("https://www.genome.jp/dbget-bin/www_bget?pathway+", kegg$Systems[i])
    kegg$Systems[i] <- text_spec(kegg$Systems[i], link = url)
  }

  reactome <- load_files(metric_table, 'Type', 'patient_coincidence_with_clusters_and_gene_systems_reactome_0.05', TRUE)
  for (i in 1:length(reactome$Systems)){
    url <- paste0("https://reactome.org/content/detail/", reactome$Systems[i])
    reactome$Systems[i] <- text_spec(reactome$Systems[i], link = url)
  }


  systems <- rbind(go, kegg, reactome)
  
  
                    
all_clusters <- group_cluster_information(hpo, systems)

for(i in 1:length(all_clusters)){
    list <- all_clusters[[i]]
    cluster_name <- names(all_clusters[i])
    cat(cluster_name, "\n")
    for(i in list){
      print(knitr::kable(i, format="markdown"))
      cat("\n")
    }
  }


    


```